---
title: "Example"
author: "Catalina Gonzalez Gomez"
date: "2/26/2024"
output: html_document
vignette: >
  %\VignetteIndexEntry{formats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
This is a example to generate a pair of connected signatures using the package cosimu. This example is divided in two sections:
- first the generation of the primary signature, that will be use as "base";
- second, the generation of the secondary signature, that will be simulated using different transitions for each layer. 

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  fig.width=8,
  fig.height=10,
  fig.fullwidth=TRUE
)
```

```{r}
library(cosimu)
set.seed(123)
```
# 1 - Primary signature
## 1.a - Modality layer 
We are going to generate signatures of 1000 genes. The primary signature will have 5% of up-regulated genes and 5% of down-regulated genes. 
```{r}
nb_ent <- 1000
p_up <- 0.05
p_down <- 0.05

M1 <- PrimaryModalityObj$new(nb_ent = nb_ent, p_up = p_up, p_down = p_down)
```

## 1.b - Submodality layer
```{r}
prop_sm_up <- c(0.5,0.3,0.2)
prop_sm_down <- c(0.5,0.3,0.2)

S1 <- PrimarySubmodalityObj$new(mod_obj = M1, prop_sm_up = prop_sm_up, prop_sm_down = prop_sm_down)
```

## 1.c - Probability layer
```{r}
P1 <- PrimaryProbabilityObj$new(nb_ent = nb_ent)
```

## 1.d - Log Fold-Change vector
### Quantile functions
```{r}
shapes <- c(4,7,12)
scale <- 0.3
qf_vect_up <- purrr::map(
  shapes,
  function(shape) {
    local(function(x) {
      p <- parent.env(environment())
      p$shape <- shape
      p$scale_r <- scale
      qgamma(x,shape,scale = scale)
    })
    }
  )
qf_vect_nr <- list(
  function(x) qnorm(x,0,0.15)
  )
qf_vect_down <- purrr::map(
  shapes,
  function(shape) {
    local(function(x) {
      p <- parent.env(environment())
      p$shape <- shape
      p$scale <- scale
      -qgamma(x,shape,scale = scale)
    })
    }
  )
qf_list <- list(qf_vect_up, qf_vect_nr, qf_vect_down)
```

### Log fold-change
```{r}
mod_vect <- M1$get_values()
submod_vect <- S1$get_values()
perc_vect <- P1$get_values()

X1 <- purrr::map_dbl(seq(1, nb_ent), function(i) {
  m <- mod_vect[i]
  sm <- submod_vect[i]
  p <- perc_vect[i]
  qf <- qf_list[[m]][[sm]]
  return(qf(p))
  })
```

## 1.e - Signature object
There is not need to do generate independently each one of the layer the `cosimu::PrimarySignatureObj` embeds this three layers and the calculation of the log fold-change vector to simplify the simulation.
```{r}
set.seed(100)
primary_sig <- PrimarySignatureObj$new(nb_ent = nb_ent, p_up = p_up, p_down = p_down,     prop_sm_up = prop_sm_up, prop_sm_down = prop_sm_down, qf_vect_up = qf_vect_up, qf_vect_nr = qf_vect_nr, qf_vect_down = qf_vect_down)
```

# 2- Secondary Signature
The same decomposition is possible for the secondary signature, but here we only present the embedded version. First user must define de parametrization for the transition between each layer.

## 2.a- Modality transition
```{r}
mod_transition <- "default"
connectivity_score <- 0.9
nr_noise <- 0.02
```

## 2.b- Sub-modality transition
```{r}
submod_transition <- "cop"
copula_submod <- "Frank"
rho_submod <- 0.9
eps_submod <- 1e-3
optim_method_submod <- "Brent"
```

## 2.c- Probability transition
```{r}
proba_transition <- "cop"
copula_prob <- "Frank"
theta_prob <- 10
```

## 2.d - Signature object
Here we use the same quantile functions than the ones used for the primary object, but it is possible to define a new ones for the secondary one, the only conditions applied for the `deterministic` and `stochastic` sub-modalities transitions are that the we should have the same number of sub-modalities between the deregulated modalities and equal to those used in the primary object.
```{r}
set.seed(200)
secondary_sig <- SecondarySignatureObj$new(
  prim_sig_obj = primary_sig, mod_transition = mod_transition,
  connectivity_score = connectivity_score, nr_noise = nr_noise,
  submod_transition = submod_transition, prop_sm_up = prop_sm_up,
  prop_sm_down = prop_sm_down, 
  rho_submod = rho_submod, copula_submod = copula_submod,
  eps_submod = eps_submod, optim_method_submod = optim_method_submod,
  proba_transition = proba_transition, theta_prob = theta_prob,
  copula_prob = copula_prob,
  qf_vect_up = qf_vect_up, qf_vect_nr = qf_vect_nr, qf_vect_down = qf_vect_down)
```

# 3- Estimation of the connectivity
Simple estimation of the connectivity score, between the pair of signatures, using Pearson's correlation score over the top 100 most up and down regulated genes of the primary signature.
```{r}
X1 <- primary_sig$get_lfc_vect()
X2 <- secondary_sig$get_lfc_vect()
up <- order(X1,decreasing = TRUE)[1:100]
down <- order(-X1, decreasing = TRUE)[1:100]
cor(X1[c(up,down)],X2[c(up,down)])
```
